---
title: "More Iteration in `R`"
author: "Jake Rozran"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include = FALSE}
library(tidyverse)
library(Lahman)

data("Teams")
Teams
options(scipen = 999)
```

# Looping `across` Columns

Last class we messed with a loop that checked if something was a numeric, 
then we calculated the average. 

Turns out, we don't need a loop for that!

```{r across}
Teams %>% 
    summarise(across(where(is.numeric), mean, na.rm = TRUE))

# Code from last class:
for(col in names(Teams)) {
    if(!is.numeric(Teams[, col])) {
        next
    }
    
    this_mean <- mean(Teams[, col], na.rm = TRUE)
    print(paste0("the mean of ", col, " is ", this_mean))
}
```

`across` also lets us specify the set of variables we want to use in 
`summarise()`.

```{r more_across}
Teams %>% 
    summarise(across(c(yearID, R:SF, BPF), mean, na.rm = TRUE))
```

# Using `map` family of functions

As we did in the loop, we can also do with the `map` family of functions to 
apply a function to each item in a list or vector. 

These functions come from `purrr` which has a 
[cheatsheet](https://github.com/rstudio/cheatsheets/blob/main/purrr.pdf)!

## Using known functions

```{r map}
Teams %>% 
    select(where(is.numeric)) %>% 
    map_dbl(mean, na.rm = TRUE)
```

We can also iterate over the rows of data, too.

```{r map_rows}
angels <- Teams %>% 
  filter(franchID == "ANA") %>% 
  group_by(teamID, name) %>%
  summarize(began = first(yearID), ended = last(yearID)) %>% 
  arrange(began)
angels

angels %>% 
    pull(name) %>% 
    map_int(nchar)
```

```{r map_regression}
mtcars %>% 
    split(.$cyl) %>%
    map(~ lm(mpg ~ wt, data = .)) %>% 
    map(summary) %>% 
    map_dbl("r.squared")
```

