---
title: "Introduction to Data Visualization"
subtitle: "STAT 4380 - 001"
author: "Jake Rozran"
institute: "Villanova University"
date: "`r Sys.Date()`"
output:
    xaringan::moon_reader:
        css: xaringan-themer.css
        nature:
            highlightStyle: github
            highlightLines: true
---

```{r setup, include = FALSE}
options(htmltools.dir.version = FALSE)
knitr::opts_chunk$set(fig.retina = 3, warning = FALSE, message = FALSE, echo = FALSE)
library(fec12)
library(dplyr)
library(tidyr)
library(ggplot2)
library(mdsr)
```

```{r xaringan-themer, include = FALSE, warning = FALSE}
library(xaringanthemer)
style_duo_accent(primary_color = "#035AA6", 
                 secondary_color = "#03A696",
                 code_font_size = "0.5rem")
```

```{r data_calc, include = FALSE, warning = FALSE}
indiv <- read_all_individuals()
contributions <- read_all_contributions()

candidate_lkup <- candidates %>% 
    select(cand_id,
           name = cand_name,
           party = cand_pty_affiliation,
           state = cand_office_st,
           district = cand_office_district,
           office = cand_office)

spent <- contributions %>% 
    group_by(cand_id, phase = transaction_pgi) %>% 
    summarise(N = n(),
              total = sum(transaction_amt),
              avg = mean(transaction_amt)) %>% 
    left_join(candidate_lkup, by = c("cand_id"))

spent2 <- contributions %>%
    mutate(spent_against = ifelse(transaction_tp %in% c('24A', '24N'), 
                                  transaction_amt, 0),
           spent_for = ifelse(!transaction_tp %in% c('24A', '24N'), 
                              transaction_amt, 0)) %>%
    group_by(cand_id, phase = transaction_pgi) %>%
    summarize(N = n(),
        total = sum(transaction_amt),
        avg = mean(transaction_amt),
        supporting = sum(spent_for),
        against = sum(spent_against)) %>%
    left_join(candidate_lkup, by = c("cand_id")) %>%
    mutate(attack_pct = against / total)

donations <- indiv %>% 
    filter(cmte_id %in% c("C00431171", "C00431445", "C00518282", "C00494740")) %>% 
    mutate(candidate = if_else(cmte_id %in% c("C00431171", "C00518282"), 
                               "Romney", "Obama")) %>% 
    filter(transaction_pgi > "") %>% 
    mutate(phase = if_else(grepl("G", transaction_pgi), "General", "Primary")) %>% 
    filter(transaction_amt > 0 & transaction_amt < 100000)
```

class: inverse, center, middle

# The 2012 Federal Election Cycle

#To Begin, a Case Study!

We are going to use an example from the book to show how important choices in
data visualization effect our understanding of what is happening.    

*Ethics Note: When done intentionally, this is how one may "lie with statistics."*

---

class: left, middle

# The Federal Election Commission (FEC) Maintains a Website of All Contributions

On this website there are: 
1. All contributions of $200 or more made by individuals to candidates and 
committees.
2. Spending by committees on behalf (and against) candidates.
3. In fact, this data is available in `R` in the `fec12` library!

---

class: center, middle

# If you had this Data and Were Tasked with an Analysis, What Questions Might you Ask?

---

class: center, middle

# A Good Place to Start: "On whose candidacy was the most money spent?"

---

class: top, left

### What Conclusions do you Draw From the Data?

```{r basic_spending, out.width = "100%", fig.height = 4.5}
first_plot <- spent %>%
    filter(total > 4000000) %>%
    mutate(spent = total / 1000000) %>%
    ggplot(aes(x = name, y = spent)) + 
        geom_col(fill = "navy") + 
        scale_x_discrete(name = NULL) + 
        scale_y_continuous(name = "Money Spent (millions of USD)", 
                           labels = scales::dollar) + 
        theme(axis.text.y = element_text(size = rel(0.5), color = "black", hjust = 1),
              axis.text.x = element_text(color = "black"),
              panel.background = element_blank(),
              axis.ticks = element_blank(),
              panel.grid.major.y = element_blank(),
              panel.grid.minor.y = element_blank(),
              panel.grid.major.x = element_line(color = "dark gray"),
              panel.grid.minor.x = element_line(color = "light gray")) +
        coord_flip()

first_plot
```

---

class: top, left

### Fun (and Obvious Fact): Committees are not Limited to Spending Money in Support of a Candidate

They Can also Spend Money **against** a candidate. Now what conclusions do you 
draw? This is the same exact data as before, with more information provided.

```{r for_against, out.width = "100%", fig.height = 3.75}
second_plot <- spent2 %>%
    filter(total > 4000000) %>% 
    select(cand_id, name, supporting, against) %>%
    pivot_longer(-c(cand_id, name), names_to = "type", values_to = "spent") %>%
    mutate(spent = spent / 1000000) %>%
    ggplot(aes(x = name, y = spent, fill = type)) + 
        geom_col() + 
        scale_fill_manual(values = c("navy", "light blue")) + 
        scale_y_continuous(name = "Money Spent (millions of USD)", 
                           labels = scales::dollar) + 
        theme(axis.text.y = element_text(size = rel(0.4), color = "black", hjust = 1),
              axis.text.x = element_text(color = "black"),
              panel.background = element_blank(),
              axis.ticks = element_blank(),
              panel.grid.major.y = element_blank(),
              panel.grid.minor.y = element_blank(),
              panel.grid.major.x = element_line(color = "dark gray"),
              panel.grid.minor.x = element_line(color = "light gray"),
              axis.title.y = element_blank(), 
              legend.title = element_blank()) +
        coord_flip()

second_plot
```

---

class: top, left

### Another Question to Explore: Was Mitt Supported by a Few Rich Donors While Barack Supported "by the People"?

- If we look at the whole election, what conclusion would you draw? 
- What else might we ask of this data?

```{r spending_hist, out.width = "100%", fig.height = 3.5}
ggplot(data = donations, aes(x = transaction_amt)) + 
    geom_histogram(fill = "navy", color = "white") + 
    scale_x_continuous(name = "Amount of Donation (USD)", 
                       labels = scales::dollar) + 
    scale_y_continuous(name = "Number of Donations",
                       labels = scales::comma) + 
    scale_color_discrete(name = NULL) + 
    facet_wrap(~ candidate) + 
    theme(panel.background = element_blank(), 
          panel.border = element_rect(fill = NA, color = "black"), 
          panel.grid.major = element_line(color = "light gray"),
          panel.grid.minor = element_blank(),
          axis.ticks = element_line(color = "light gray"),
          strip.background = element_blank(), 
          strip.text = element_text(size = 12))
```

---

class: top, left

### Diving Deeper: Let's Change the Data and Again Answer the Same Question

Was Mitt Supported by a Few Rich Donors While Barack Supported "by the People"?

```{r split_donations, out.width = "100%", fig.height = 4}
split_donies <- ggplot(data = donations, aes(x = transaction_amt, color = candidate)) + 
    geom_density(adjust = 4) + 
    scale_x_log10(name = "Amount of Donation (USD)", labels = scales::dollar) + 
    scale_color_manual(values = c("navy", "light blue")) +
    facet_wrap(~ phase) + 
    theme(panel.background = element_blank(), 
          panel.border = element_rect(fill = NA, color = "black"), 
          panel.grid.major = element_line(color = "light gray"),
          panel.grid.minor = element_blank(),
          axis.ticks = element_line(color = "light gray"),
          strip.background = element_blank(), 
          strip.text = element_text(size = 12), 
          legend.title = element_blank(), 
          legend.position = "top")

split_donies
```

---

class: inverse, middle, center

# Composing Data Graphics

Anyone can create a data visualization - the hard part is making it tell the 
correct story in the easiest to understand way. 

```{r bad_good}
bad_dat <- data.frame(variable = c("a", "b", "c", "d", "e"),
                      value = c(22, 21, 20, 19, 18))

bad_graph <- ggplot(bad_dat, aes(x = "", y = value, fill = variable)) +
    geom_bar(stat = "identity") + 
    coord_polar("y", start = 0) + 
    theme(legend.position = "top",
          legend.title = element_blank(),
          panel.background = element_blank(),
          axis.ticks = element_blank(),
          axis.title = element_blank())

better_graph <- ggplot(bad_dat, aes(x = variable, y = value, fill = variable)) +
    geom_bar(stat = "identity") + 
    geom_text(aes(label = value), color = "black", nudge_y = 1) + 
    theme(legend.position = "top",
          legend.title = element_blank(),
          panel.background = element_blank(),
          axis.ticks = element_blank(),
          axis.title = element_blank(),
          panel.grid.major.y = element_line(color = "light gray"))
```

--

.pull-left[
```{r bad_graph, out.width = "100%", fig.height = 6}
bad_graph
```
]

--

.pull-right[
```{r better_graph, out.width = "100%", fig.height = 6}
better_graph
```
]

---

class: middle, center

# To the Point of the Last Slide, it has Taken me Hours to the Previous Slides Just how I Want Them

It is _easy_ to slap something together. It is **HARD** to put something clean, 
clear, and meaningful together.

---

class: top, left

# A taxonomy for data graphics

Taxonomy, smaxonomy. That's just a fancy name for saying you got to know the 
finer parts of a data viz before you, too, can make fine data viz. 

.pull-left[
Data graphics can be understood in terms of four basic elements: 

1. Visual cues
2. Coordinate systems
3. Scale
4. Context
]

.pull-right[
*And two bonus items:* 

1. Facets
2. Layers
]

---

class: top, left

# Visual Cues

- Visual cues are graphical elements that draw the eye to what we want our 
audience to focus upon
- These are the building blocks of data viz
- Human beings' ability to perceive difference in magnitude accurately decends 
in this order
    - For example, humans are not as good at perceiving angles as we are length, 
    hence why the example before with the pie chart was so difficult
    - Also, humans are poor at perceiving differences in color, that is why 
    being thoughtful about color is so important (and why heatmaps are 
    frequently poopooed)

---

class: top, left

# Visual Cues (Continued)

```{r viz_cue}
tribble(
        ~`Visual Cue`, ~`Variable Type`, ~Question,
        "Position", "numerical", "where in relation to other things?",
        "Length", "numerical", "how big (in one dimension)?",
        "Angle", "numerical", "how wide? parallel to something else?",
        "Direction", "numerical", "at what slope? in a time series, going up or down?",
        "Shape", "categorical", "belonging to which group?",
        "Area", "numerical", "how big (in two dimensions)?",
        "Volume", "numerical", "how big (in three dimensions)?",
        "Shade", "either", "to what extent? how severely?",
        "Color", "either", "to what extent? how severely?"
    ) %>%
        mdsr_table(caption = "Visual cues and what they signify.") %>%
        kableExtra::column_spec(3, width = "24em")
```

---

class: top, left

### Position, Length, & Area Example

Because I feel like this is getting a little too much me talking at you, let us
do an example together. 

Someone tell me what they see in this plot, especially as it relates to 
**position, length, and area**?

.pull-left[
```{r position_code, echo = TRUE, fig.show = "hide", out.width = "100%", fig.height = 5}
plot_dat <- iris %>%
    group_by(Species) %>% 
    summarise(avg_sepal_length = mean(Sepal.Length))

ggplot(plot_dat, aes(x = Species, y = avg_sepal_length)) + 
    geom_bar(stat = "identity")
```
]

--

.pull-right[
```{r ref.label = "position_code", echo = FALSE, out.width = "100%", fig.height = 5}
```
]

---

class: top, left

### Direction

What **direction** do we see in this plot? Anything else?

.pull-left[
```{r direction_code, echo = TRUE, fig.show = "hide"}
ggplot(iris, aes(x = Sepal.Length, y = Sepal.Width)) + 
    geom_point()
```
]

--

.pull-right[
```{r ref.label = "direction_code", echo = FALSE, out.width = "100%", fig.height = 5}
```
]

---

class: top, left

### Angle, Color, Shape

Talk to me about the **angle, color, and shape**.

.pull-left[
```{r acs_code, echo = TRUE, fig.show = "hide"}
ggplot(iris, aes(x = Sepal.Length, 
                 y = Sepal.Width, 
                 color = Species, 
                 shape = Species)) + 
    geom_point() + 
    geom_smooth(method = "lm", se = FALSE)
```
]

--

.pull-right[
```{r ref.label = "acs_code", echo = FALSE, out.width = "100%", fig.height = 5}
```
]

---

### *Angle, Color, Shape* and Size

We've just added another dimension: **size**. What do we realize now?

.pull-left[
```{r size_code, echo = TRUE, fig.show = "hide"}
ggplot(iris, aes(x = Sepal.Length, 
                 y = Sepal.Width, 
                 color = Species, 
                 shape = Species,
                 size = Petal.Length)) + 
    geom_point() + 
    geom_smooth(method = "lm", se = FALSE)
```
]

--

.pull-right[
```{r ref.label = "size_code", echo = FALSE, out.width = "100%", fig.height = 5}
```
]

