---
title: "Working with Data - Part II"
author: "Jake Rozran"
date: "`r Sys.Date()`"
output: html_document
---

# Matricies

A matrix is basically a 2 dimensional vector (all a single type of variable)

```{r create_matrix}
x <- c(5, 7, 9, 13, -4, 8) # Our vector from last lecture

A <- matrix(x, 2, 3) # Creates a matrix from x
A 
```

You can also see that a matrix is also a vector.

```{r checkout_the_matrix}
is.matrix(A) # Is A a matrix?
is.vector(A) # Is A a vector?
is.matrix(x) # Is x a matrix?
is.vector(x) # Is x a vector?
is.numeric(x) # Is x a numeric?
is.numeric(A) # Is A a numeric?
```

We can find items in the matrix, the same as we can with a vector. The 
difference is that we have to provide both a row and a column (in that order).

```{r find_matrix_items}
A[2, 3] # Give me the item in the 2nd row and 3rd column
A[, 1] # Give me the items in the 1st column
A[1, ] # Give me the items in the 1st row
```

# Data Frames

Data Frames are similar to matricies: 2 dimensional. They will 
support different types of data, though. 

```{r df}
y <- c(20, 13, 17, 92, 1, 4) # Our other vector from last class

ds <- data.frame(x, y) # Create a data frame with x and y
ds
```

We can do some basic things to see summary stats of a data frame.

```{r df_functions}
dim(ds) # What are the dimensions of ds?
str(ds) # What is the structure of ds?
summary(ds) # What is the summary of ds?
```

Create a new data frame with a few different data types in it

```{r diff_df}
new_df <- data.frame(name = c("Jacob", "Jingle", "Heimer", "Schmidt"),
                     age = c(14, 22, 100, 9),
                     is_old = c(FALSE, FALSE, TRUE, FALSE))

dim(new_df) # What are the dimensions of new_df?
str(new_df) # What is the structure of new_df?
summary(new_df) # What is the summary of new_df?
```

There are also a few different ways we can find things in a data frame:

```{r find_df}
new_df[2, 3] # Can find items by row and column number
new_df[, c("name", "age")] # Can find items by column names
new_df[1, c("name", "age")] # Can find items by column names and row number
new_df[1, "name"] # Can find items by column names and row number
new_df$is_old[2] # Can specify the column name behind the $ and then the row number
new_df$is_old # Can specify the column name behind the $ without the row number
```

Let's create a (nonsensical) bigger data frame to see how head and tail work. 
We can use the rbind command (which just binds a data frame by rows, or 
basically stacks data frames on top of each other).

```{r head_tail}
another_df <- rbind(new_df, 
                    new_df,
                    new_df,
                    new_df,
                    new_df)
head(another_df) # lets us see the first 6 rows (by default)
?head
head(another_df, n = 5) # lets us see the first 5 rows
tail(another_df) # lets us see the last 6 rows (by default)
```


# Testing & Changing Object Types

- `is.` is a prefect function that creates a logical result if the type matches.
    - `any()` tests all of an object.
- `as.` is a prefix function that changes the type of an object (temporarily).

```{r testing}
is.na(c(1, 2, 3, NA))

any(is.na(c(1, 2, 3, NA)))

gender <- c("male", "female", "male", "female")
is.factor(gender)
is.character(gender)

summary(as.factor(c(1,2,2,1,3,1,1,2)))
```

# If Then Statements

There are two types of if statements in `R`.

## `if`, `else if`, `else`

I use these types of if statements to direct the program in a direction. `R` 
calls it a "control-flow statement". IRL, I have an if statement to make one 
type of variable if the data is monthly and another type of variable if the 
data is daily and another type if the data is weekly.

```{r if_statement}
type <- "monthly"

if (type == "monthly") {
    variable <- "this is monthly data so I do a monthly calculation"
} else if (type == "weekly") {
    variable <- paste0("this is weekly data so I do a weekly calculation (whi",
                       "ch may be totally different then monthly)")
} else {
    variable <- "this must be daily data (since that is the only other option)"
}

print(variable)
# Change the type and see how the output changes
```

## `ifelse` or `if_else`

`ifelse` and the tidyverse version `if_else` (you must first load the package 
`tidyverse` to use it) are used to evaluate an entire vector, item by item, and 
do something. I use these functions when I need to make a calculation based on 
a vector. 

For example, how many females are there in our data from before? 

```{r ifelse}
sum(ifelse(gender == "female", 1, 0))

# LETS LOOK UNDER THE HOOD
ifelse(gender == "female", 1, 0)
```

We can do the same, again, with the `tidyverse` version.

```{r if_else}
library(tidyverse)

sum(if_else(gender == "female", 1, 0))

# LETS LOOK UNDER THE HOOD
if_else(gender == "female", 1, 0)

```

# IN CLASS ASSIGNMENT 

Pause here for the in-class assignment
