---
title: "Working with Data - Part 1"
author: "Jake Rozran"
date: "`r Sys.Date()`"
output: 
    html_document:
        toc: true
        toc_float:
            smooth_scroll: true
            collapsed: false
        
---

# Assignment

```{r assignment}
x <- c(5, 7, 9, 13, -4, 8) # preferred
x
x = c(5, 7, 9, 13, -4, 8) # equivalent
print(x) # Equivalent to just typing x

c(5, 7, 9, 13, -4, 8) -> x # also works
x
```

The above code creates a vector of length 6 using the c() function to 
concatenate scalars.

*Note: `c(5, 7, 9, 13, -4, 8) = x` will NOT work.*

# Useful Commands for Data Entry

```{r data_entry}
60:72 # Integers from 60 to 72
seq(60, 72) # Also integers from 60 to 72

seq(60, 72, by = 2) # Also integers from 60 to 72, but by 2
seq(60, 70, length = 3) # Three equally spaced values

rep(60, 5) # Repeat the value 60, 5 times
rep(c(60, 72), 3) # Repeat the sequence c(60, 72) 3 times.

rep(c(60, 66, 72), c(1, 2, 3)) # One 60, two 66s, and three 72s
```

# Missing Data in R

- Missing data in R is denoted by `NA`.
- Some functions produce `NA` when any data are missing

```{r nas}
mean(c(1, 2, NA))
```

If a function has the `na.rm` argument, you can use that (preferred) to get rid 
of pesky `NA`s - IRL, you better understand why there are `NA`s in your data, or 
else you might be calculating something unintended!

Also, if there is no `na.rm` argument, you can leverage the function `na.omit()`.

```{r no_more_nas}
mean(c(1, 2, NA), na.rm = TRUE) # Preferred for mean function

mean(na.omit(c(1, 2, NA))) # Also works
```

# Data Classes

There are six basic data types: `logical`, `character`, `numeric` (which can be 
further broken down into `integer`, `double`, and `complex`), and `raw.` 

## Factors

There are also factors, but they are annoying, so we won't use them much

```{r factors}
z <- factor(c("x", "y", "z"))
z
z <- append(z, c("a", "b", "c"))
z
```

This happens because factors are stored internally as integers that correspond 
to the id’s of the factor levels. So we get the integer of the "level" instead 
of the actual character value. 

This is just much easier to deal with and gives the expected result:

```{r char}
z <- c("x", "y", "z")
z <- append(z, c("a", "b", "c"))
z
```

# Vectors 

## Data Selection for Vectors

Since vector operations are so fundamental in R, it is important to be able to 
access (or index) elements within these vectors. Many different ways of indexing 
vectors are available.

```{r vectors}
x <- c(5, 7, 9, 13, -4, 8)

x[2] # Gets the second item from the vector x
x[c(2, 4)] # Gets the second and fourth item from the vector x
x[c(TRUE, TRUE, TRUE, TRUE, TRUE, FALSE)] # Only returns those values that are TRUE
x[1:5] # Returns the first through the fifth values of x
x[-6] # Returns x without the sixth value
```

Vectors are recycled if needed; for example, when comparing each of the elements 
of a vector to a scalar.

```{r vec_ops}
x > 8 # Greater than
x < 8 # Less than
x == 7 # Equal
x != 7 # Not equal
x <= 5 # Less than or equal to
x >= 5 # Greater than or equal to
```

## Side Note about `TRUE` and `FALSE`

ALSO - `TRUE` == 1 and `FALSE` == 0

```{r sum_true}
sum(x > 8) # Count of how many elements meet the condition x > 8
```

## Operators & Vectors

Operators can also be done on a vector

```{r vec_arth}
x + 1
x - 1
x * 4
x / 2
x^3
x %% 6
x %/% 5
```

You can also do these operations on two vectors.

```{r more_vec}
y <- c(20, 13, 17, 92, 1, 4)

x + y
x^y
x - y
x * y
```

## The %in% command

```{r in_cmd}
"c" %in% c("a", "b", "c", "d", "e", "f")
"z" %in% c("a", "b", "c", "d", "e", "f")
```

