---
title: "Intro R - Working with Data"
output:
  word_document: default
---

Enter the following data and assign it to a variable called `prediet`: 140, 160, 
135, 120, 125, 165, 108, 130, 225, 140, 180.

```{r}
prediet <- c(140, 160, 135, 120, 125, 165, 108, 130, 225, 140, 180)
prediet
```


Enter the following data and assign it to a variable called `postdiet`: 139, 
155, 120, 118, 123, 155, 108, 129, 195, 142, 176

```{r}
postdiet <- c(139, 155, 120, 118, 123, 155, 108, 129, 195, 142, 176)
```

Calculate the weight loss (from `prediet` to `postdiet`).

```{r}
weight_loss <- postdiet - prediet
weight_loss
```

Print (only) the 5th element of weight loss using `[]` notation.

```{r}
weight_loss[5]
```

Convert this value to kilograms (2.2 pounds is a kilogram)

```{r}
kgs <- weight_loss / 2.2
kgs
```

Verify that the kilograms is a numeric variable using `class()`.

```{r}
class(kgs)
is.numeric(kgs)
```

Assemble all four variables into a data frame using `data.frame()`.

```{r}
weight <- data.frame(prediet, postdiet, weight_loss, kgs)
weight
```


Find the find the `mean`, `median`, `minimum`, and `maximum` of the weight loss 
in kilograms using a summary statistic function for each one.

```{r}
mean(weight_loss)
median(weight_loss)
min(weight_loss)
max(weight_loss)
```


Use the `summary()` command to get all the summary statistics for the weight loss 
variable in kilograms. (add a chunk)

```{r summary}
summary(weight_loss)
```

Use an `if…else` statement to create a new variable called `weightgain` which 
equals “lost” if they lost weight and “gain” otherwise. Verify that it’s a 
character variable by using the `is.character()` function. (add a chunk)

```{r new_var}
library(tidyverse)
weight$weightgain <- if_else(weight_loss < 0, "lost", "gain")
weight
is.character(weight$weightgain)
```

Conduct the appropriate statistical inference to determine whether the weight 
loss program was effective. State the assumptions/conditions for this test and 
verify them. (We haven't yet talked about graphics, but you can use the `stem()` 
command to do a basic text-based graphic or try `hist()`. (add a chunk)

```{r stats}
hist(weight_loss)

quantile(weight_loss, 0.95)
```

H0: weight_loss >= 0
HA: weight_loss < 0













