---
title: "Simulation"
author: "Jake Rozran"
date: '`r Sys.Date()`'
output: html_document
---

```{r setup, include = FALSE}
library(tidyverse)
library(furrr)

options(scipen = 999)
```

## Monte Carlo Simulation

What is a monte carlo simulation - write that out here.

### Simple Example

```{r monte_carlo, warning = FALSE}
get_rnorm <- function(nsamp) {
    tmp <- tibble(mean = mean(rnorm(nsamp)),
                  samp_size = nsamp)
    return(tmp)
}

test <- 1:100 %>% 
    map_dfr(~get_rnorm(10))

ggplot(test, aes(x = mean)) + 
    geom_bar()
```

### Real World Example

Let's talk about being a successful mutual fund manager. A fund manager comes to 
you and says, "I've been successful for the past 10 years straight. I'm a genius. 
Give me $10MM and I'll make you a fortune!" What is your reaction?

The cool thing is that we can test this out and see for ourselves, using a 
Monte Carlo simulation, how "skilled" this person really is!

What if we assume that in any given year, irrespective of the year you had 
before, you have a 10% chance at being a successful mutual fund manager. It's 
hard and in other words, one year's success is independent of the next year's 
success. Let's also assume that there are a million mutual fund managers on day 
1.

Is our person lucky? Or did they do something that required skill?

```{r fund_mgr, warning = FALSE}
fund_success <- function(num_years = 10, success_pct = 0.5) {
    get_success <- tibble(
        year = 1:num_years,
        success = ifelse(runif(num_years) <= success_pct, TRUE, FALSE)
    )
    
    years_in_biz <- min(get_success$year[get_success$success == FALSE])
    
    if(is.infinite(years_in_biz)) {
        years_in_biz <- 10
    }
    
    return(years_in_biz)
}

num_mgr <- 10000

successful_fund_mgr <- 1:num_mgr %>% 
    map_dbl(~fund_success()) 

plot_dat <- tibble(mgr = 1:num_mgr,
                   years_successful = successful_fund_mgr)

ggplot(plot_dat, aes(x = successful_fund_mgr)) + 
    geom_bar()
```

Let's speed things up a bit with parallel processing. 

```{r future, warning = FALSE}
future::plan(multisession)

num_mgr <- 100000

successful_fund_mgr <- 1:num_mgr %>% 
    future_map_dbl(~fund_success(), 
                   .progress = TRUE,
                   .seed = TRUE) 

plot_dat <- tibble(mgr = 1:num_mgr,
                   years_successful = successful_fund_mgr)

ggplot(plot_dat, aes(x = successful_fund_mgr)) + 
    geom_bar()

table(plot_dat$years_successful)
```


This doesn't seem like much, right? You look at Wall Street... stocked full of 
men (mainly men) dressed in high priced suits and driving expensive cars. Surely,
this can't be the whole picture. 

We'll... this is only one year. PLUS, we don't see the ones that wash out. They 
move away or work at the restaraunts that all the Wall St. guys go to. Let's see 
how this plays out over 10 years.


```{r years_10, warning = FALSE}
years <- 30
num_mgr <- 100000

successful_fund_mgr <- 1:(num_mgr * years) %>% 
    future_map_dbl(~fund_success(), 
                   .progress = TRUE,
                   .seed = TRUE) 

plot_dat <- tibble(year = rep(1:years, each = num_mgr),
                   mgr = rep(1:num_mgr, years),
                   years_successful = successful_fund_mgr) %>% 
    mutate(final_year = year + years_successful - 1,
           years_successful = as.character(years_successful))
plot_dat

ggplot(plot_dat, aes(x = final_year, 
                     group = factor(years_successful, levels = 10:1),
                     fill = factor(years_successful, levels = 10:1))) + 
    geom_bar() + 
    theme(legend.title = element_blank(),
          legend.position = "top")

ggplot(plot_dat[plot_dat$final_year <= 30, ], 
       aes(x = final_year, 
           group = factor(years_successful, levels = 10:1),
           fill = factor(years_successful, levels = 10:1))) + 
    geom_bar() + 
    theme(legend.title = element_blank(),
          legend.position = "top")
```
