---
title: "Tidy Data and Reshaping Data"
author: "Jake Rozran"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include = FALSE}
library(tidyverse)
```

# What is Tidy Data?

First, let's install and load a new package, which lets us read data from 
google sheets directly.

```{r googlesheets}
# install.packages("googlesheets4")
library(googlesheets4)
```

Then - let's grab a dataset from Gapminder (this key is in 
[Google Drive](https://docs.google.com/document/d/1zq1jd9SoViPaTgA31uwvTk3Sg3UeUieWIaTdDjhYSAE/edit)).

```{r gapminder}
hiv_key <- "1kWH_xdJDM4SMfT_Kzpkk-1yuxWChfurZuWYjfmv51EA"
hiv <- read_sheet(hiv_key)
hiv
```

This data is NOT tidy. Why is it not tidy? 

Each row has multiple datapoints or observations.

It is "easy" to see the trend in this data... just following a row. However, 
each row does not refer to a specific, unique thing. 

So... what data is tidy?

```{r babynames}
# install.packages("babynames")
library(babynames)
data("babynames")
babynames
```

Each row is a name, year, gender, and then `n` number of people who were born 
and named that. This is "tidy."

Why is it tidy, again? For one, we can easily figure out the top names.

```{r babynames_popular}
babynames %>% 
    group_by(sex, name) %>% 
    summarise(total_births = sum(n)) %>% 
    arrange(desc(total_births))
```

We could do this by `year`, too... 

```{r babynames_popular_by_year}
babynames %>% 
    filter(year == max(babynames$year)) %>% 
    group_by(sex, name, year) %>% 
    summarise(total_births = sum(n)) %>% 
    arrange(desc(total_births))
```

And `gender`... 

```{r babynames_popular_by_year_and_sex}
babynames %>% 
    filter(year == max(babynames$year) & 
               sex == "M") %>% 
    group_by(sex, name, year) %>% 
    summarise(total_births = sum(n)) %>% 
    arrange(desc(total_births))
```

So... yes... this is tidy. 

# Reshaping Data 

All is not lost, though, if our data is not tidy. 

In comes `pivot_wider` and `pivot_longer` to help us go to and from tidy format. 

*PS - earlier I taught `reshape2` to some of you. That is a valid option, too, *
*but not what the book teaches. Whatever works, works.*

## Wide to long

```{r pivot_longer}
hiv_long <- hiv %>% 
    select(`Estimated HIV Prevalence% - (Ages 15-49)`:`2008`) %>%
    pivot_longer(-`Estimated HIV Prevalence% - (Ages 15-49)`, 
                 names_to = "year", 
                 values_to = "percent") %>% 
    na.omit() %>% 
    mutate(year = as.numeric(year),
           percent = percent / 100)
hiv_long
```

Before we made this change, we couldn't plot the data. Now we can!

```{r plot_the_long}
ggplot(hiv_long, aes(x = year, y = percent, 
                     group = `Estimated HIV Prevalence% - (Ages 15-49)`,
                     color = `Estimated HIV Prevalence% - (Ages 15-49)`)) + 
    geom_line() + 
    theme(legend.position = "none") 
```

## Pivot Wider

We can also go the other way, for any number of reasons. Let's say we are 
working with a colleague who relies on Excel? They will need the data in wide 
format to graph it or use it.

```{r pivot_wider}
baby_wide <- babynames %>% 
    filter(name %in% c("Jacob", "Lucas", "Elizabeth", "Regina", "Sasha", "Poppy")) %>% 
    select(year:n) %>% 
    pivot_wider(names_from = year, values_from = n)
baby_wide
```

