---
title: "Two Table Wrangling - In Class"
author: "your name here"
date: "2/24/2022"
output: html_document
---

```{r setup, include = FALSE}
library(Lahman) # INSTALL THIS PACKAGE IF NECESSARY
library(tidyverse)
```

## Manny Ramirez

1. There is a dataset called `Batting` in the `Lahman` package. Start by 
creating a dataset named `manny` that filters Manny Ramirez's data from the 
`Batting` data. His `playerID` is `ramirma02`.

```{r 1}
manny <- Batting %>% 
    filter(playerID == "ramirma02")
manny
```

2. From this data, create a summary that calculates Manny's career:
    - Number of years played
    - First year played
    - Last year played
    - Batting Average (total hits (`H`) divided by total at bats (`AB`))
    - Total Homeruns (`HR`)
    - Total Runs Batted In (`RBI`)
    
```{r 2}
manny_summary <- manny %>% 
    summarise(years_played = max(yearID) - min(yearID),
              first_year = min(yearID),
              last_year = max(yearID),
              batting_avg = sum(H) / sum(AB),
              total_HR = sum(HR),
              total_RBI = sum(RBI))
manny_summary
```
    
3. Create a new variable `manny_by_year` that calculates by year the following: 
    - Batting Average (total hits (`H`) divided by total at bats (`AB`))
    - Total Homeruns (`HR`)
    - Total Runs Batted In (`RBI`)
    - Number of teams he played on
    - MAKE SURE THAT YOU KEEP MANNY'S `playerID` AS PART OF THIS DATASET
    
```{r manny_by_year}
manny_by_year <- manny %>% 
    group_by(yearID) %>%
    summarise(total_ba = sum(H) / sum(AB),
              total_hr = sum(HR),
              total_rbi = sum(RBI),
              total_teams = length(teamID)) %>% 
    mutate(playerID = "ramirma02")
manny_by_year
```

4. There is another dataset called `Master`, which includes details on 
each player in the dataset. Join the `Master` dataset to the `manny_by_year` 
dataset and then create a new column that calculates Manny's age for that 
season. The output should only include `playerID`, `yearID`, age during that 
year, batting average, homeruns, runs batted in, and the number of teams he 
played on. 

```{r age_question}
manny_age <- manny_by_year %>%
    inner_join(Master, by = c("playerID" = "playerID")) %>%
    mutate(age = yearID - birthYear) %>% 
    select(playerID,
           yearID,
           age,
           total_ba,
           total_hr,
           total_rbi,
           total_teams)

manny_age
```


5. Create a new variable `cal_by_year` that calculates the following by year for 
Cal Ripken Jr. (`ripkeca01`): 
    - Batting Average (total hits (`H`) divided by total at bats (`AB`))
    - Total Homeruns (`HR`)
    - Total Runs Batted In (`RBI`)
    - Number of teams he played on
    
```{r cal_by_year}
cal <- Batting %>% 
    filter(playerID == "ripkeca01")

cal_by_year <- cal %>% 
    group_by(yearID) %>% 
    summarise(total_ba = sum(H) / sum(AB),
              total_hr = sum(HR),
              total_rbi = sum(RBI),
              total_teams = length(teamID)) %>% 
    mutate(playerID = "ripkeca01")

cal_by_year <- Batting %>% 
    filter(playerID == "ripkeca01") %>% 
    group_by(yearID) %>% 
    summarise(total_ba = sum(H) / sum(AB),
              total_hr = sum(HR),
              total_rbi = sum(RBI),
              total_teams = length(teamID)) %>% 
    mutate(playerID = "ripkeca01")
cal_by_year
```
    
6. Using an `inner_join`, join the `manny_by_year` and `cal_by_year` datasets. 
Which years did their careers overlap? Who was better (according to these 
metrics)?

```{r manny_and_cal}
new_data <- manny_by_year %>% 
    inner_join(cal_by_year, by = c("yearID" = "yearID"))

new_data %>% 
    select(yearID,
           total_rbi.x,
           total_rbi.y)
```










