---
title: "Data Visualization - Part 2"
output: html_document
---

```{r setup}
library(tidyverse)
```

Let's take a look at the MPG dataset again. This dataset contains a subset of the 
fuel economy data that the EPA makes available on http://fueleconomy.gov. It 
contains only models which had a new release every year between 1999 and 2008 - 
this was used as a proxy for the popularity of the car.

Explore the data to see what's in there...

```{r explore}
data("mpg")

head(mpg)
str(mpg)
summary(mpg)
```

Run the code to make a graph

```{r run_this_graph}
ggplot(data = mpg, aes(x = displ, y = hwy)) +
    geom_point()
```

Change the `x` and `y` axes to be logarithmic (both independently and then at 
the same time). Which seems to best fit the data?

```{r change_the_axes}
ggplot(data = mpg, aes(x = displ, y = hwy)) +
    geom_point()
```

If you used `scale_y_continuous` in the last part, change it to `coord_trans` 
for each example (if you did `coord_trans`, use the appropriate `scale_` this 
time). Which do you prefer?

```{r}
ggplot(mpg, aes(x = class, y = hwy)) + 
    geom_point()
```

Change the color of the points to be be controlled by `drv`.

```{r}

```

Move the legend from the last plot to the top of your graph.

```{r}

```

Instead of color, can you create a facet by `drv`?

```{r}

```

From the last plot, add in an appropriate title, `x` label, and `y` label.

```{r}

```



