---
title: "Data Science - Homework #3"
author: "your name"
output:
 word_document: default
---



<!-- 
- Change your name above and save the file. You should get in the habit of 
doing this for future assignments.
- Recall that the setup option always runs this section once before others.
- I included the necessary libraries in the setup chunk below. If you use install.packages, make sure to comment it out after you execute it once. Better 
yet, go ahead and delete it. I know that if your library ran, you must have 
installed the package.
- I added `message = FALSE` and `warning = FALSE` as well. These stifle printing 
messages and warnings, respectively, to the Knitted file. Feel free to 
experiment with those and other chunk commands, if you want.
- You should delete unnecessary text, but please leave the problem headers for 
each problem (the ones in **bold**)

- AFTER EACH GRAPH YOU CREATE, MAKE SURE TO COMMENT ON WHAT YOU SEE
- PLEASE ADD APPROPRIATE LABELS TO EACH GRAPH
--> 


```{r setup, message=F, warning=F}
library(tidyverse)
library(mosaicData) # For problem 2
library(babynames) # For problem 3
library(mdsr) # For problem 4
```

## Problem 1) Arrests Per State in the US

The `USArrests` dataset contains one observation per US State. The variables are 
the murder, assault, and rape rates per 100,000 people as well as the percent of 
the population living in urban areas.

**Problem 1a) Create a graph of murder rate. **

```{r}

```

Comment on what you see (feel free to erase this line)



**Problem 1b) Create a graph of assault rate. **

```{r}

```

Comment on what you see (feel free to erase this line)




**Problem 1c) Create a scatterplot of murder rate (`x`) vs. assault rate (`y`). **

```{r}

```

Comment on what you see (feel free to erase this line)


**Problem 1d) Add a `geom_smooth` to the previous plot. **


```{r}

```


Comment on what you see (feel free to erase this line)


**Problem 1e) Add another `geom_smooth` with a `method = "lm"` argument to the **
**previous plot. (You might want to make it a different color to distinguish it **
**from the previous one.) **

```{r}

```

Comment on whether the linear model is appropriate for these data (feel free to 
erase this line)




**Problem 1d) Add `UrbanPop` as a color aesthetic to the scatterplot you created **
**in 1c) above. **


```{r}

```


Comment on what you see (feel free to erase this line)






## Problem 2) Galton Height Data

The Galton data can be found in the `mosaicData` package. The data contains 
`height` information on all children within a `family`, including the 
child's `sex` (gender), the `mother`'s height, the `father`'s height, and 
the number of kids (`nkids`) in the family. Height is measured in inches.


**Problem 2a) Create a scatterplot of each person's height against their **
**Father's height. (As typically done, please put the independent variable on **
**the x-axis)**

```{r}


```

Comment on what you see (feel free to erase this line)


**Problem 2b) Add a regression line to the previous plot by using** 
**`geom_smooth(method = "lm")`. **

```{r}


```

Comment on what you see (feel free to erase this line)


**Problem 2c) Add a y = x line to the previous plot using ** 
**`geom_abline(intercept = 0,slope = 1)` **

```{r}


```

Comment on what you see (feel free to erase this line)



**Problem 2d) Add `sex` as an aesthethic to the previous plot. Make sure you **
**have two separate regression lines, one per each `sex`. **


```{r}


```

Comment on what you see (feel free to erase this line)




