---
title: "Data Science - Homework #3"
author: "Jake Rozran"
output: html_document
---



<!-- 
- Change your name above and save the file. You should get in the habit of 
doing this for future assignments.
- Recall that the setup option always runs this section once before others.
- I included the necessary libraries in the setup chunk below. If you use install.packages, make sure to comment it out after you execute it once. Better 
yet, go ahead and delete it. I know that if your library ran, you must have 
installed the package.
- I added `message = FALSE` and `warning = FALSE` as well. These stifle printing 
messages and warnings, respectively, to the Knitted file. Feel free to 
experiment with those and other chunk commands, if you want.
- You should delete unnecessary text, but please leave the problem headers for 
each problem (the ones in **bold**)

- AFTER EACH GRAPH YOU CREATE, MAKE SURE TO COMMENT ON WHAT YOU SEE
- PLEASE ADD APPROPRIATE LABELS TO EACH GRAPH
--> 


```{r setup, message = FALSE, warning = FALSE}
library(tidyverse)
library(mosaicData) # For problem 2
library(babynames) # For problem 3
library(mdsr) # For problem 4
```

## Problem 1) Arrests Per State in the US

The `USArrests` dataset contains one observation per US State. The variables are 
the murder, assault, and rape rates per 100,000 people as well as the percent of 
the population living in urban areas.

**Problem 1a) Create a graph of murder rate. **

```{r 1a}
data("USArrests")
head(USArrests)

ggplot(USArrests, aes(x = Murder)) +
    geom_histogram(bins = 15)
```

Unimodal, skewed right, no outliers.



**Problem 1b) Create a graph of assault rate. **

```{r 1b}
ggplot(USArrests, aes(x = Assault)) +
    geom_histogram(bins = 15)
```

The plot is almost uniform with two modes (so... I guess we can say it is 
bimodal). No clear skew. No clear outliers. 

**Problem 1c) Create a scatterplot of murder rate (`x`) vs. assault rate (`y`). **

```{r 1c}
ggplot(USArrests, aes(x = Murder, y = Assault)) +
    geom_point()
```

There is a clear positive correlation between Murder and Assault Rates. Strong, 
positive, linear relationship between the two variables. No clear outliers. 


**Problem 1d) Add a `geom_smooth` to the previous plot. **


```{r 1d}
ggplot(USArrests, aes(x = Murder, y = Assault)) +
    geom_point() +
    geom_smooth()
```


Adding the `geom_smooth` gives a generally linear regression line to the plot. 
It does seem, however, that the assault rates have an upper boundary of sorts, 
even as the murder rate increases. 


**Problem 1e) Add another `geom_smooth` with a `method = "lm"` argument to the **
**previous plot. (You might want to make it a different color to distinguish it **
**from the previous one.) **

```{r 1e}
ggplot(USArrests, aes(x = Murder, y = Assault)) +
    geom_point() +
    geom_smooth() + 
    geom_smooth(method = "lm", color = "green")
```

A linear model does seem to be appropriate for these data. Aside from the 
deviation as murder rates reach their maximum, the lines are nearly identical. 
It is hard to tell from this graph, but I don't think there is a true deviation 
from linear there... 


**Problem 1f) Add `UrbanPop` as a color aesthetic to the scatterplot you created **
**in 1c) above. **


```{r 1f}
ggplot(USArrests, aes(x = Murder, y = Assault, color = UrbanPop)) +
    geom_point()
```

It isn't exactly easy to tell, but there seems to be a pretty random spread of 
urban population in the graph. I don't see a clear correlation between urban
population and either variable.


## Problem 2) Galton Height Data

The Galton data can be found in the `mosaicData` package. The data contains 
`height` information on all children within a `family`, including the 
child's `sex` (gender), the `mother`'s height, the `father`'s height, and 
the number of kids (`nkids`) in the family. Height is measured in inches.


**Problem 2a) Create a scatterplot of each person's height against their **
**Father's height. (As typically done, please put the independent variable on **
**the x-axis)**

```{r 2a}
data("Galton")
head(Galton)

ggplot(Galton, aes(x = father, y = height)) + 
    geom_point()
```

There seems to be a positive, moderate strength, linear relationship between a 
father's height and the child's height. 


**Problem 2b) Add a regression line to the previous plot by using** 
**`geom_smooth(method = "lm")`. **

```{r 2b}
ggplot(Galton, aes(x = father, y = height)) + 
    geom_point() +
    geom_smooth(method = "lm")

```

The linear model fits ok. There relationship is not very strong, so there is 
quite a bit of spread from the lm. That said, the spread is rather equal. 


**Problem 2c) Add a y = x line to the previous plot using ** 
**`geom_abline(intercept = 0, slope = 1)` **

```{r 2c}
ggplot(Galton, aes(x = father, y = height)) + 
    geom_point() +
    geom_smooth(method = "lm") +
    geom_abline(intercept = 0, slope = 1)

```

There regression line is certainly much flatter than a = 1. 

**Problem 2d) Add `sex` as an aesthethic to the previous plot. Make sure you **
**have two separate regression lines, one per each `sex`. **


```{r 2d}
ggplot(Galton, aes(x = father, y = height, color = sex)) + 
    geom_point() +
    geom_smooth(method = "lm") +
    geom_abline(intercept = 0, slope = 1)
```

It is quite remarkable how the lines are parallel to one another. The difference 
is about 5 inches? 




