---
title: "Intro to RMarkdown"
author: "Jake Rozran"
date: "`r Sys.Date()`"
output: 
    html_document:
        toc: true
        toc_float:
            smooth_scroll: true
            collapsed: true
---

# How to write `.md`

This is just regular paragraph text.  
And this is more plain text.

*this is italics* OR _italics_  
**this is bold** OR __bold__  
x^2^  
~~this text will be struck out~~  

[I want to hyperlink this line.](www.google.com)

We talked about making things `look like code`.  
But also, we can do R code like this `r 1 + 1`.  

```
here is some code
and another line of bullshit code
```

# Header 1

## Header 2

### Header 3

#### Header 4

##### Header 5

###### Header 6

$A = \pi*r^{2}$ 

***  

* unordered list
* item 2
    + sub-item 1
    + sub-item 2

- item 1
    - item 2
    - item 3
- item 4
    - item 5

1. ordered list
2. item 2
    a. sub-item 1
    b. sub-item 2

Table Header  | Second Header | this is a third column
------------- | ------------- | ----------------------
Table Cell    | Cell 2 | cell 6
Cell 3        | Cell 4 | 27

```
Plain text  
End a line with two spaces to start a new paragraph.  
*italics* and _italics_  
**bold** and __bold__  
superscript^2^  
~~strikethrough~~  
[link](www.rstudio.com)  
`inline code`  
\```
code chunk
with lots of lines
\```  

# Header 1

## Header 2

### Header 3

#### Header 4

##### Header 5

###### Header 6

endash: --  
emdash: ---  
ellipsis: ...  
inline equation: $A = \pi*r^{2}$  

horizontal rule (or slide break):  

***  

> block quote  

* unordered list
* item 2
    + sub-item 1
    + sub-item 2

1. ordered list
2. item 2
    + sub-item 1
    + sub-item 2

Table Header  | Second Header
------------- | -------------
Table Cell    | Cell 2
Cell 3        | Cell 4
```

Here is some more inline R code. `r sqrt(10)`.

```{r}
this <- "is text "
that <- "is more text"

text <- paste0(this, that)

print(text)
```

```{r you_name_chunk, echo = FALSE}
hist(rnorm(1000))
```


```{r another_chunk, message = FALSE}
library(tidyverse)
```




















