---
title: "Data Wrangling Project"
author: "Jake Rozran"
date: "`r Sys.Date()`"
output: html_document
---

```{r twitter_part, include = FALSE}
library(rtweet)

token <- readRDS(Sys.getenv("TWITTER_PAT"))

twitter_data <- get_timelines("hamsterofthe", n = 1000, token = token)
twitter_data[, 1:16]
twitter_data <- twitter_data %>% 
    mutate(text = gsub("ðŸ¹", "HAMSTER NAME:", text)) %>% 
    mutate(text = gsub("ðŸ·ï¸", "TAGGED USER:", text)) %>%
    mutate(text = gsub("ï¸ðŸ·", "TAGGED USER:", text)) %>%
    mutate(text = gsub("ðŸ’•", "HEARTS", text)) 
# write.csv(twitter_data[, 1:16], "hamster_data.csv", row.names = FALSE)
```

```{r setup, include=FALSE}
library(tidyverse)
library(lubridate)

twitter_data <- read_csv("hamster_data.csv")
twitter_data
```

# TEXT WRANGLING

For the following questions, use `grepl()` function. This function can find 
patterns in text.

1. Are all of the hamsters named?
    a. Create a new column called `hname` searches for "NAME" in the `text` 
    column and creates the value `TRUE` if it is there and `FALSE` if not
    b. Create a summary variable named `hamster_name` that sums and sorts the 
    count each value in `hname` from highest count to lowest count
    c. Output `hamster_name`
    d. How many posts do not include the name of the hamster?
    
```{r problem1}
# PART A
twitter_data <- twitter_data %>%
    mutate(hname = grepl("NAME", text))

# PART B
hamster_name <- twitter_data %>% 
    group_by(hname) %>% 
    summarise(count = length(hname)) %>% 
    arrange(desc(count))

# PART C
hamster_name

# PART D
twitter_data %>% 
    filter(hname == FALSE) %>% 
    pull(hname) %>% 
    length()
```

2. How many tagged users come from instagram? (hint: use the technique from 
the last problem)

```{r problem2}
# PART A
twitter_data <- twitter_data %>%
    mutate(instagram = grepl("instagram", tolower(text)))

# PART B
instagram <- twitter_data %>% 
    group_by(instagram) %>% 
    summarise(count = length(hname)) %>% 
    arrange(desc(count))

# PART C
instagram

# PART D
twitter_data %>% 
    filter(instagram == TRUE) %>% 
    pull(instagram) %>% 
    length()
```

3. Create a histogram of favorite counts.
    a. Is the standard amount of bins appropriate? If not, assign the 
    appropriate number of bins.
    b. Make the outline of the bars black and the color of the bars navy
    c. Ensure that the background of the graph is clear (or white)
    d. Remove any tick marks
    e. Remove any vertical grid lines. The horizontal grid lines should be light 
    gray.
    f. Make sure the axes are labeled nicely
    g. Make sure there is a title on the plot
    h. Describe the distribution

```{r problem3}
ggplot(twitter_data, aes(x = favorite_count)) + 
    geom_histogram(color = "black",
                   fill = "navy") + 
    xlab("Favorite Count") + 
    ylab("Count") + 
    ggtitle("Histogram of Favorite Count of Hamster of the Day") + 
    theme(panel.background = element_blank(),
          panel.grid.major.x = element_blank(),
          panel.grid.minor.x = element_blank(),
          panel.grid.major.y = element_line(color = "light gray"),
          panel.grid.minor.y = element_line(color = "light gray"),
          axis.ticks = element_blank())
```

The plot is bimodal (one mode at 0 and the other around 275), skewed right, and 
there are no outliers.

4. Which hamster got the most favorites?

```{r problem4}
twitter_data %>% 
    filter(favorite_count == max(twitter_data$favorite_count)) %>% 
    pull(text)
```

5. Create a histogram of retweet counts.
    a. Is the standard amount of bins appropriate? If not, assign the 
    appropriate number of bins.
    b. Make the outline of the bars white and the color of the bars purple
    c. Ensure that the background of the graph is light gray
    d. Change the color of any tick marks to purple
    e. Remove any vertical grid lines. The horizontal grid lines should be white.
    f. Make sure the axes are labeled nicely
    g. Make sure there is a title on the plot
    h. Describe the distribution

```{r problem5}
ggplot(twitter_data, aes(x = retweet_count)) + 
    geom_histogram(color = "white",
                   fill = "purple") + 
    xlab("Favorite Count") + 
    ylab("Count") + 
    ggtitle("Histogram of Favorite Count of Hamster of the Day") + 
    theme(panel.background = element_rect(fill = "light gray"),
          panel.grid.major.x = element_blank(),
          panel.grid.minor.x = element_blank(),
          panel.grid.major.y = element_line(color = "white"),
          panel.grid.minor.y = element_line(color = "white"),
          axis.ticks = element_line(color = "purple"))
```

The plot is bimodal (one mode at 0 and the other around 50), skewed right, and 
there may be outliers above 150.

6. Which hamster got the most retweets?

```{r problem6}
twitter_data %>% 
    filter(retweet_count == max(twitter_data$retweet_count)) %>% 
    pull(text)
```

7. Is there a correlation between age of tweet (in days) and favorite count? You 
will need to create a new column that counts the age of the tweet. I am asking 
if the longer a tweet is up, does it get more likes? Or is it just a standard 1 
day of visibility or something like that? (hint: remember the `lubridate` 
package)

```{r problem7}
twitter_data <- twitter_data %>% 
    mutate(age_of_tweet = Sys.Date() - as_date(created_at))

ggplot(twitter_data, aes(x = age_of_tweet, y = favorite_count)) + 
    geom_point() + 
    geom_smooth()
```

Really hard to tell what is happening here! The short answer is... no, there 
does not seem to be a trend that the older tweets get more favorites.

8. What is the first day in the dataset? What is the last day in the dataset?

```{r problem8}
twitter_data %>% 
    summarise(first_day = min(created_at),
              last_day = max(created_at))
```

10. Are there any days with multiple posts? How many days?

```{r problem10}
twitter_data %>% 
    group_by(as_date(created_at)) %>% 
    summarise(count = length(text)) %>% 
    filter(count > 1) %>% 
    arrange(desc(count))

twitter_data %>% 
    group_by(as_date(created_at)) %>% 
    summarise(count = length(text)) %>% 
    filter(count > 1) %>% 
    arrange(desc(count)) %>% 
    dim()
```

12. Read in the weather data. 
    a. What are the first and last day in the dataset?
    b. How many weather stations are there?
    c. Do all the stations have the same amount of entries?
    d. Which station has the most entries?

```{r problem12}
weather_data <- read_csv("weather_data.csv")

## PART A
weather_data %>% 
    summarise(first_day = min(DATE), 
              last_day = max(DATE))

# PART B
weather_data %>% 
    pull(NAME) %>% 
    unique() %>% 
    length()

# PART C
weather_data %>% 
    group_by(NAME) %>% 
    summarise(count = length(NAME))

# PART D
weather_data %>% 
    group_by(NAME) %>% 
    summarise(count = length(NAME)) %>% 
    slice_max(count, n = 1)
```




12. Join the hamster data to the weather data for the top station from the 
previous problem (if there are any ties, pick your favorite). What issues do you 
have and what choices did you make to join this data together?

```{r problem12}
join_weather_data <- weather_data %>% 
    filter(NAME == "PHILADELPHIA INTERNATIONAL AIRPORT, PA US")

twitter_data <- twitter_data %>% 
    mutate(join_date = as_date(created_at))

joined_data <- twitter_data %>% 
    left_join(join_weather_data, by = c("join_date" = "DATE"))
```

Chose to use left join and had to make a date that matched the date from weather. 

13. Plot the daily average temperature from the last 30 days of the dataset. 
Create a data label that adds the favorite count of the daily hamster to the 
plot (if there are multiple, you may choose one or paste all together).

```{r problem13}
plot_data <- joined_data %>% 
    arrange(desc(join_date)) %>% 
    head(n = 30) %>% 
    group_by(join_date) %>% 
    summarise(avg_temp = mean(TAVG),
              label = paste0(favorite_count, collapse = ","))

ggplot(plot_data, aes(x = join_date, y = avg_temp)) + 
    geom_line() + 
    geom_text(aes(label = label), nudge_y = 1)
```

14. What is the average precipitation (`PRCP`) at "PHILADELPHIA INTERNATIONAL 
AIRPORT, PA US"?

```{r problem14}
weather_data %>% 
    filter(NAME == "PHILADELPHIA INTERNATIONAL AIRPORT, PA US") %>% 
    summarise(avg_prcp = mean(PRCP, na.rm = TRUE))
```

15. What is the overall average (`TAVG`), minimum (`TMIN`), and maximum (`TMAX`) 
temperature of this data at "PHILADELPHIA INTERNATIONAL AIRPORT, PA US"?

```{r problem15}
weather_data %>% 
    filter(NAME == "PHILADELPHIA INTERNATIONAL AIRPORT, PA US") %>% 
    summarise(avg_temp = mean(TAVG, na.rm = TRUE),
              avg_min = mean(TMIN, na.rm = TRUE),
              avg_max = mean(TMAX, na.rm = TRUE))
```

16. What is the average, minimum, and maximum temperature from this data by 
month at "PHILADELPHIA INTERNATIONAL AIRPORT, PA US"?

```{r problem16}
weather_data %>% 
    filter(NAME == "PHILADELPHIA INTERNATIONAL AIRPORT, PA US") %>% 
    group_by(month(DATE)) %>% 
    summarise(avg_temp = mean(TAVG, na.rm = TRUE),
              avg_min = mean(TMIN, na.rm = TRUE),
              avg_max = mean(TMAX, na.rm = TRUE))
```

17. What is the average, minimum, and maximum from this data by year at 
"PHILADELPHIA INTERNATIONAL AIRPORT, PA US"?

```{r problem17}
weather_data %>% 
    filter(NAME == "PHILADELPHIA INTERNATIONAL AIRPORT, PA US") %>% 
    group_by(year(DATE)) %>% 
    summarise(avg_temp = mean(TAVG, na.rm = TRUE),
              avg_min = mean(TMIN, na.rm = TRUE),
              avg_max = mean(TMAX, na.rm = TRUE))
```
18. What is the average, minimum, and maximum from this data by year and month 
at "PHILADELPHIA INTERNATIONAL AIRPORT, PA US"?

```{r problem18}
weather_data %>% 
    filter(NAME == "PHILADELPHIA INTERNATIONAL AIRPORT, PA US") %>% 
    group_by(month(DATE), year(DATE)) %>% 
    summarise(avg_temp = mean(TAVG, na.rm = TRUE),
              avg_min = mean(TMIN, na.rm = TRUE),
              avg_max = mean(TMAX, na.rm = TRUE))
```
19. Plot the average daily temperature for the entire dataset grouped by station. 
What issues do you see with the plot? Make it the clearest possible. 

