---
title: "Intro to `ggplot2`"
author: "Jake Rozran"
date: "`r Sys.Date()`"
output: 
    html_document:
        toc: true
        toc_float:
            smooth_scrolling: true
            collapse: false
---

```{r setup, include = FALSE}
library(mdsr)
library(tidyverse)

options(scipen = 999)
```

# The Data We'll Use for Class Today

```{r data_intro}
data("CIACountries") # GIMME THE DATA

?CIACountries # WHAT IS THIS DATA?

head(CIACountries) # SHOW ME THE DATA
str(CIACountries) # SHOW ME THE DATA DIFFERENTLY
summary(CIACountries) # SUMMARIZE THE DATA FOR ME

unique(CIACountries$country) # WHAT COUNTRIES ARE HERE?
```

# The `ggplot` object

- `ggplot2` is a library to build data graphics
- The reason that `ggplot2` has been so successful is that it allows you to build 
graphics incrementally
- The first thing we do is create a `ggplot` object
- There are two methods for doing this (I prefer the second, as it makes it 
clearer at a glance what all the aesthetics and layers are)

```{r intro}
# Book Method - Assign a variable the ggplot object
# Then use that variable to build upon
g <- ggplot(data = CIACountries, aes(x = educ, y = gdp))
g + geom_point()

# I prefer to keep the object unassigned so I can see all the parts
# This becomes especially useful as you build your graphics over huge scripts
ggplot(data = CIACountries, aes(x = educ, y = gdp)) +
    geom_point()
```

## What are we Looking at Here?

1. The `ggplot` object
    - This informs which data to use - in this case `CIACountries` and then 
    which aethestics to use with the `aes()` function
    - Here, we are telling `ggplot` to use the `educ` column for the `x` 
    position and `gdp` for the `y` position
2. The `geom_point()` layer
    - This is the "glyph" we want to plot
    - Each point is a different country, located at its `educ` and `gdp`
    - *Note:* the size for each point is the same, so it is not an aesthetic 
    (it does not map to a visual cue)
    
## Adding More Visual Cues

- You can add additional aesthetics beyond `x` and `y`.
- You can add them in the `ggplot` object (if they will be used universally) OR
- You can add them in the layer object

```{r getting_more}
ggplot(data = CIACountries, aes(x = educ, y = gdp)) +
    geom_point(aes(color = net_users))
```

### To Change an Aestetic or Not, That is the Question

You can also customize a layer universally (not based on a variable in the data)

In the code below, we change the size of the dot to be bigger for all dots 
(still not a visual cue, but does make the plot easier to read).

```{r non_aes_change}
ggplot(data = CIACountries, aes(x = educ, y = gdp)) +
    geom_point(aes(color = net_users), size = 3)
```

We can, however, also make size an aesthetic (and add even more information to 
the plot)!

*Note: There is a such a thing as too much information in a plot... be careful *
*not to add too many aesthetics.*

```{r aes_change}
ggplot(data = CIACountries, aes(x = educ, y = gdp)) +
    geom_point(aes(color = net_users, size = roadways))
```

### Let's get Wild - there is a World of Things to Change

- All of these "features" can be assigned via `aes()` or not (which makes them 
universal changes)
- `alpha` is the one that may seem less intuitive, it is the opacity (between 0 
and 1)
- `color` (or `colour`, if you prefer) is the color of lines
- `fill` is the color filling an object (like a bar or the squares below)
- There are a bunch of shapes built in 
([I googled it and there is a list](https://www.datanovia.com/en/blog/ggplot-point-shapes-best-tips/))

```{r all_the_options}
ggplot(data = CIACountries, aes(x = educ, y = gdp)) +
    geom_point(size = 3, 
               color = "purple", 
               alpha = 0.5, 
               shape = 22, 
               fill = "pink")
```

### Going Beyond Points

- Obviously there are more than just scatter plots in this world
- You can build any plot you can imagine with `ggplot2` (I may not have a great 
imagination)
- Check out the `ggplot2` cheat sheet on the 
[RStudio Website](https://www.rstudio.com/resources/cheatsheets/) or our 
[Google Drive](https://drive.google.com/drive/u/2/folders/1UDfF9IGHay82mZo97NBvgkKoEjSIfP39)

```{r add_text}
ggplot(data = CIACountries, aes(x = educ, y = gdp)) +
    geom_text(aes(label = country, color = net_users))
```

### The Beat and the Chorus at the SAME TIME!

We're not really ready for this yet, but we can add multiple layers, too!

```{r text_and_points}
ggplot(data = CIACountries, aes(x = educ, y = gdp, color = net_users)) +
    geom_point() +
    geom_text(aes(label = country), 
              nudge_x = 0.15, 
              check_overlap = TRUE, 
              hjust = 0)
```

# IN CLASS EXERCISE!


