---
title: "In Class with `dplyr`"
author: "Jake Rozran"
date: "2/15/2022"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(lubridate)
options(scipen = 999)

stock_data <- read_csv("~/Documents/rle_nightly/data/nightly_run/pick_tracker_new.csv")
stock_data
```

1. Select the following columns: `symbol`, `test_start_date`, `test_end_date`,
`start`, `purchase_price`, `purchase_amount`, `current_price`, `security_type`

```{r select}
stock_data %>% 
    select(symbol, test_start_date, test_end_date, start, purchase_price,
           purchased_amount, current_price, security_type)
```

2. Filter for Crypto and `start` of "64 day pred"

```{r filter}
stock_data %>% 
    filter(security_type == "Crypto" & 
               start == "64 day pred")
```

3. create a column that pulls out the year of the start date (use lubridate 
`year` function)

```{r create_column}
stock_data %>% 
    mutate(test_start_date = as_date(test_start_date, format = "%m/%d/%y"),
           start_year = year(test_start_date)) %>% 
    select(symbol, test_start_date, start_year)
```

4. Create a count of how many purchases there were for each `start` value

```{r purchase_count}
stock_data %>% 
    group_by(start) %>% 
    summarise(purchase_count = length(symbol))
```

5. Rename the columns `test_start_date` and `test_end_date` to `start_date` and
`end_date`, respectively

```{r rename}
stock_data %>%
    rename(start_date = test_start_date, 
           end_date = test_end_date) %>%
    select(symbol, 
           start_date,
           end_date)
```

6. Which stock has the most rows?

```{r which_is_most}
stock_data %>% 
    group_by(symbol) %>% 
    summarise(symbol_count = length(symbol)) %>%
    arrange(desc(symbol_count)) %>% 
    head(1)
```















