---
title: "Scraping Web Data in R"
author: "Jake Rozran"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(rvest)
library(lubridate)
```

## Scraping Web Data

What if you want to grab data directly from a webpage? It is in a table there, 
but... for whatever reason, you can't pull it directly from a different source. 
Wikipedia is a great example of this. 

Let's look at [earthquakes from 1905](https://en.wikipedia.org/wiki/List_of_earthquakes_in_1905).

There are a bunch of tables here... I want to analyze it myself. Enter the 
`rvest` package.

```{r webscraping}
url <- "https://en.wikipedia.org/wiki/List_of_earthquakes_in_1905"

tables <- url %>% 
    read_html() %>% 
    html_nodes("table")
tables

tables %>% 
    pluck(3) %>% 
    html_table()
```

Rad, right? Let's get each monthly table (why didn't they put these in one 
table? who knows... )

```{r putting_tables_together}
earthquakes <- data.frame()
for(table in c(6, 8, 10, 12, 14, 16, 18, 20, 22, 24, 26, 28)) {
    tmp_tab <- url %>%
        read_html() %>% 
        html_nodes("table") %>% 
        pluck(table) %>% 
        html_table()
    
    earthquakes <- bind_rows(earthquakes, tmp_tab)
}

earthquakes
```

So... that worked, kind of. We have all the data but there is some weirdness. 

Who can name some of the weirdness?

Let's do it again and fix it as we go?

```{r putting_tables_together_fully}
earthquakes <- data.frame()

table_dict <- data.frame(table_num = c(6, 8, 10, 12, 14, 16, 18, 20, 22, 24, 26, 
                                       28),
                         month = c(1:12),
                         year = 1905)

for(table in table_dict$table_num) {
    tmp_tab <- url %>%
        read_html() %>% 
        html_nodes("table") %>% 
        pluck(table) %>% 
        html_table()
    
    names(tmp_tab) <- as.character(tmp_tab[1, ])
    tmp_tab <- tmp_tab %>% 
        mutate(Date = gsub("\\[\\d+\\]", "", Date),
               Date = as_date(paste0(table_dict$year[table_dict$table_num == table],
                                   "-", 
                                   table_dict$month[table_dict$table_num == table],
                                   "-", Date)),
               Mw = as.numeric(Mw),
               `Depth (km)` = as.numeric(`Depth (km)`),
               Dead = as.numeric(Dead),
               Injured = as.numeric(Injured)) %>% 
        filter(Date <= as_date("1906-01-01")) 
    
    earthquakes <- bind_rows(earthquakes, tmp_tab)
}

earthquakes
```

So, that's rad... who can think of another topic where we can pull some data 
from? 