---
title: "Joining Tables - Continued"
author: "Jake Rozran"
date: '`r Sys.Date()`'
output: html_document
---

```{r setup, include = FALSE}
library(tidyverse)
```

## Remember Joins?

From last class, we know we can do a bunch of types of joins:

1. `inner_join`: get back only the rows that match from the 2 datasets
2. `left_join`: get back all of the rows from the left dataset, regardless if 
they have a match on the right dataset
3. `right_join`: get back all of the rows from the right dataset, regardless if 
they have a match on the left dataset
4. `full_join`: get back all of the rows from both datasets, regardless if there 
is match in the other dataset.

## Join Examples

```{r data}
left_data <- data.frame(id = 1:10,
                        type = "left dataset")
left_data

right_data <- data.frame(id = 6:15,
                         class = "right dataset")
right_data
```

```{r inner_join}
inner <- left_data %>% 
    inner_join(right_data, by = c("id" = "id"))
inner
```

```{r left_join}
left <- left_data %>% 
    left_join(right_data, by = c("id" = "id"))
left
```

```{r right_join}
right <- left_data %>% 
    right_join(right_data, by = c("id" = "id"))
right
```

```{r full_join}
full <- left_data %>% 
    full_join(right_data, by = c("id" = "id"))
full
```

```{r anti_join}
anti_left <- left_data %>% 
    anti_join(right_data, by = c("id" = "id"))
anti_left

anti_right <- right_data %>% 
    anti_join(left_data, by = c("id" = "id"))
anti_right
```