---
title: "Joins AGAIN!"
author: "Jake Rozran"
date: '2022-03-08'
output: html_document
---

```{r setup, include = FALSE}
library(tidyverse)
```

## Joins

Two weeks ago, we learned about joins. We are taking two different datasets and 
joining them on a shared column.

1. `inner_join`: return back to us a dataset with only those rows that match 
from both datasets. 
2. `left_join`: return back all of the rows from the "left" dataset, regardless 
if they match anything from the "right" dataset
3. `right_join`: return back all of the rows from the "right" dataset, 
regardless if they match anything from the "left" dataset
4. `full_join`: returns back all rows from both datasets regardless of match.

```{r data}
left_data <- data.frame(id = 1:10,
                        type = "left dataset")
left_data

right_data <- data.frame(id = 6:15,
                         class = "right dataset")
right_data
```

```{r inner_join}
inner <- left_data %>% 
    inner_join(right_data, by = c("id" = "id"))
inner
```

```{r left_join}
left <- left_data %>% 
    left_join(right_data, by = c("id" = "id"))
left
```

```{r right_join}
right <- left_data %>% 
    right_join(right_data, by = c("id" = "id"))
right
```

```{r full_join}
full <- left_data %>% 
    full_join(right_data, by = c("id" = "id"))
full
```










