---
title: "Iteration in Class Exercise"
author: "Jake Rozran"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include = FALSE}
library(tidyverse)
library(mosaicData)
library(nycflights13)
```

1. Use the `HELPrct` data from the `mosaicData` package to calculate the mean of 
all numeric variables (be sure to exclude missing values).

```{r problem1}
HELPrct %>% 
    summarise(across(where(is.numeric), mean, na.rm = TRUE))

HELPrct %>% 
    select(where(is.numeric)) %>% 
    map_dbl(mean, na.rm = TRUE)
```

2. You have data about flight delays in a tibble called `flights` from the 
`nycflights13` package. Write a function that, for any given airport code (e.g., 
LGA), will return a tibble with two columns, the airport code, and the average 
arrival delay time. 

```{r problem2}

delay_fnct <- function(airport_code) {
    airport_delay <- flights %>% 
        filter(dest == airport_code) %>%
        summarise(airport_code = airport_code,
                  delay = mean(arr_delay, na.rm = TRUE))
        
    return(airport_delay)
}

new_variable <- delay_fnct("PHL")
new_variable

```

3. Using the appropriate `map` function, create a data frame with every 
airport code and its average arrival delay time. Arrange the results from the 
most delayed to the least. (use the function you just created above)










