---
title: "Data Visualization - Part 1"
output:
word_document: default
---

```{r setup}
library(tidyverse)
```

Let's take a look at the MPG dataset.This dataset contains a subset of the 
fuel economy data that the EPA makes available on http://fueleconomy.gov. It 
contains only models which had a new release every year between 1999 and 2008 - 
this was used as a proxy for the popularity of the car.

Explore the data to see what's in there...

```{r explore}
data("mpg")

head(mpg)
str(mpg)
summary(mpg)
table(mpg$manufacturer)

```

Run the code to make a graph

```{r run_this_graph}
ggplot(data = mpg, aes(x = displ, y = hwy)) +
    geom_point()
```

Add `color`, `size`, `alpha`, and `shape` aesthetics to your graph. Experiment.

```{r add_some_cool_stuff}
ggplot(data = mpg, aes(x = displ, y = hwy)) +
    geom_point(aes(size = cyl,
                   color = manufacturer,
                   alpha = cty,
                   shape = class))

ggplot(data = mpg, aes(x = displ, y = hwy)) +
    geom_point(size = 3,
               color = "blue",
               alpha = 0.5,
               shape = 11)
```

Replace this scatterplot with one that draws boxplots. Use the cheatsheet. 
Try your best guess.

```{r}
ggplot(mpg, aes(x = class, y = hwy)) + 
    geom_boxplot()
```

Make the histogram of the `hwy` variable.Hint: Do not supply a `y` variable.

```{r}
ggplot(mpg, aes(x = hwy)) +
    geom_histogram(binwidth = 5)
```

Make a density plot of `hwy` colored by `class`. Use the cheatsheet and/or try 
your best guess.

```{r}
ggplot(mpg, aes(x = hwy, color = class)) +
    geom_density()
```

Make the bar chart of `class`.

```{r}
ggplot(mpg, aes(x = class)) +
    geom_bar()
```

Predict what this code will do. Then run it.

```{r}
ggplot(mpg, aes(x = displ, y = hwy)) + 
    geom_point() +
    geom_smooth()
```



