---
title: "Reshaping Data"
author: "YOUR NAME HERE"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include = FALSE}
library(tidyverse)
library(mdsr)
```

1. Is `mtcars` a tidy dataset?

```{r mtcars}
data("mtcars")
mtcars$name <- row.names(mtcars)
mtcars
```
Yes. This is tidy. Each item (row) represents the same kind of thing. 

2. Using the `mtcars` dataset, use `pivot_longer` to turn this into a long 
dataset. Is this tidy? Why or why not?

```{r pivot_longer}
longer_mtcars <- mtcars %>% 
    pivot_longer(-name, names_to = "variable", values_to = "value")
longer_mtcars
```
This is not tidy data, because each row kind of represents a different thing. 

3. Is `population` a tidy dataset?

```{r population}
data("population")
population
```

This is a tidy dataset. 

4. Using the `population` dataset, use `pivot_wider` to turn this into a wide 
dataset. Is this tidy? Why or why not?

```{r pivot_pop_wider}
pop_wide <- population %>% 
    pivot_wider(names_from = year, 
                values_from = population)
pop_wide
```
    
    
    
    
    
    
    
    
    
    
    










