---
title: "Intro R - Importing Data"
output:
  pdf_document: default
  word_document: default
  html_document:
    df_print: paged
---

Step 1: Load the `tidyverse` package.

```{r setup, include = FALSE}
library(tidyverse)
library(readxl)
```

Step2: Download the file `titanic.csv` from the datasets directory on Google 
Drive. Read it into R using `read_csv()` from the readr package. Explore the 
data to make sure it was read in properly.

```{r read_csv}
titanic <- read_csv("../../datasets/titanic.csv")
str(titanic)
head(titanic)
summary(titanic)
table(titanic$Sex)
table(titanic$Survived)
```


Step 3: Compare the file when you read it in using `read.csv()` (base `R`) 
versus `read_csv()` (`readr`). After exploring, what do you see different? 

```{r read.csv}
titanic_other <- read.csv("../../datasets/titanic.csv")
head(titanic_other)
```

Step 4: Download the file `manufacturing.xlsx` from the datasets directory on 
Google Drive.  Read it into R using `read_xlsx()` from the `readxl` package. 
Explore the data to make sure it was read in properly.

```{r readxl}
manufacturing <- read_xlsx("../../datasets/manufacturing.xlsx",
                           sheet = 2)
head(manufacturing)
manufacturing
summary(manufacturing)
table(manufacturing$Product)
table(manufacturing$Shift)
```

