---
title: "Iteration in R - A Lesson in Bad `R` Form"
author: "Jake Rozran"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include = FALSE}
library(tidyverse)
library(mdsr)
library(Lahman)
```

# Iteration in R

Many of you are likely already aware of loops, no matter your coding experience. 

There are two main types of loops in `R`:

1. `for` loops
2. `while` loops

We are going to learn loops today... then we are going to do our best to avoid 
using them for the remainder of the term. 

I create loops in my code all the time... they are intuitive to me and they 
work. However, they are very inefficient in `R` and there are some better ways 
to handle iteration (which we'll learn over the next few classes)!

## `for` loops

In a `for` loop, you tell `R` exactly what to iterate over. 

```{r basic_for_loop}
for(i in 1:6) {
    print(i)
}
```

```{r more_for}
for(team in unique(Teams$name)) {
    print(team)
}
```

All loops will repeat the code on the inside for each "iteration" defined. 

What if we want to output the number of seasons for each team in the `for` loop?

```{r even_more_for}
total_data <- data.frame()
for(team in unique(Teams$name)) {
    team_data <- Teams %>% 
        filter(name == team)
    
    first_year <- min(team_data$yearID)
    last_year <- max(team_data$yearID)
    seasons <- last_year - first_year + 1
    
    tmp <- data.frame(name = team,
                      first_season = first_year,
                      last_season = last_year,
                      seasons = seasons)
    
    total_data <- bind_rows(total_data, tmp)
}
total_data
```

There's a better way to do this, though, right?

```{r less_for}
total_data <- Teams %>% 
    group_by(name) %>%
    summarise(first_season = min(yearID),
              last_season = max(yearID)) %>% 
    mutate(seasons = last_season - first_season + 1)
total_data
```
What if we want to calculate the average for each column that is a numeric?

We could go column by column and copy/paste the code... 

```{r averages}
str(Teams)
```

```{r looping_across_colums}
col_names <- names(Teams)
for(col in col_names) {
    if(!is.numeric(Teams[, col])){
        next
    } 
    
    string <- paste0("The Average ", col, " is ", mean(Teams[, col], na.rm = TRUE))
    print(string)
}
```

We'll learn in the next class how to deal with this without the loop. 

## `while` loops

`while` loops are very similar, but will run until a condition is no longer 
`TRUE`.

```{r simple_while}
i <- 0
while(i <= 4) {
    print(i)
    i <- i + 1
}
```

These are good for semi-dynamic functionality. If you don't know your budget, 
but you know you need to repeat something until your budget is spent, you can 
use a `while` loop. 