---
title: "Loops in `R`"
author: "Jake Rozran"
date: '2022-03-22'
output: html_document
---

```{r setup, include = FALSE}
library(tidyverse)
library(Lahman)
```

# Iteration in `R`

Many of you might know about loops already. In `R` there are two types that we 
are going to talk about.

1. `for` loop
2. `while` loop

We are going to learn about these and then we are going to do our best to not 
use as we progress through the class. 

I use loops all the time. I encourage you to use them if they work. We'll learn 
other methods that are better in `R`.

## `for` loops

```{r basic_for_loop}
for(i in 1:6) {
    print(i)
}
1:6
```

```{r another_for_loop}
str(Teams)

for(team in unique(Teams$name)) {
    print(team)
}

unique(Teams$name)
```

```{r summary_for}
total_data <- data.frame()

for(team in unique(Teams$name)) {
    team_data <- Teams %>% 
        filter(name == team)
    
    first_year <- min(team_data$yearID)
    last_year <- max(team_data$yearID)
    seasons_in_existance <- last_year - first_year + 1
    
    tmp_data <- data.frame(name = team,
                           first_season = first_year,
                           last_season = last_year,
                           seasons = seasons_in_existance)
    
    total_data <- bind_rows(total_data, tmp_data)
}

total_data

team <- "Philadelphia Phillies"

team_data <- Teams %>% 
        filter(name == team)
team_data    

first_year <- min(team_data$yearID)
last_year <- max(team_data$yearID)
seasons_in_existance <- last_year - first_year + 1

tmp_data <- data.frame(name = team,
                       first_season = first_year,
                       last_season = last_year,
                       seasons = seasons_in_existance)
tmp_data

total_data <- bind_rows(total_data, tmp_data)
total_data
```

```{r the_better_way}
summary_data <- Teams %>% 
    group_by(name) %>% 
    summarise(first_season = min(yearID),
              last_season = max(yearID)) %>% 
    mutate(seasons = last_season - first_season + 1)
summary_data
```


```{r across_columns}
str(Teams)
names(Teams)

for(col in names(Teams)) {
    if(!is.numeric(Teams[, col])) {
        next
    }
    
    this_mean <- mean(Teams[, col], na.rm = TRUE)
    print(paste0("the mean of ", col, " is ", this_mean))
}
```

## `while` loops

```{r basic_while}
i <- 0
while(i <= 4) {
    print(i)
    i <- i + 1
}

```



















