---
title: "More `ggplot2`"
author: "Jake Rozran"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include = FALSE}
library(mdsr)
library(tidyverse)

options(scipen = 999)

# THE DATA WE'LL CONTINUE TO USE
data("CIACountries") # GIMME THE DATA
```

# Moving on From Aesthetics - Let's Discuss Scale

Harken back to last class. We had this plot:

```{r harken}
ggplot(data = CIACountries, aes(x = educ, y = gdp)) +
    geom_point(aes(color = net_users, size = roadways))
```

At the bottom of the plot, we have a hard time discerning differences in GDP. 
There are lots of points stacked on top or near one another.

One option we have is to adjust the scale. 

As with many things in `R`, there is more than one way to accomplish this.

## `coord_trans`

```{r log_y_coord_trans}
ggplot(data = CIACountries, aes(x = educ, y = gdp)) +
    geom_point(aes(color = net_users, size = roadways)) +
    coord_trans(y = "log10")
```

## `scale_y_continuous`

```{r log_y_scale_y}
ggplot(data = CIACountries, aes(x = educ, y = gdp)) +
    geom_point(aes(color = net_users, size = roadways)) +
    scale_y_continuous(trans = "log10")
```

## Side-by-Side

```{r side_by_side}
ct <- ggplot(data = CIACountries, aes(x = educ, y = gdp)) +
    geom_point(aes(color = net_users, size = roadways)) +
    coord_trans(y = "log10") + 
    theme(legend.position = "none")

syc <- ggplot(data = CIACountries, aes(x = educ, y = gdp)) +
    geom_point(aes(color = net_users, size = roadways)) +
    scale_y_continuous(trans = "log10") + 
    theme(legend.position = "none")

gridExtra::grid.arrange(ct, syc, nrow = 1)

```

I tend to like the plot on the left better because it is more clear that the 
y axis has been translated to a logarithmic scale. 

You are not, however, limited to only tranfering the y axis. There are similar 
functions to to perform similar operations on different aesthetics:

- `scale_x_continuous()`
- `scale_x_discrete()`
- `scale_y_discrete()`
- `scale_color()`
- etc. 

