---
title: "Importing Data"
author: "Jake Rozran"
date: "`r Sys.Date()`"
output: 
    html_document:
        toc: true
        toc_float:
            smooth_scrolling: true
---

## Read it in as a vector using `c()`.

```{r vector}
height <- c(59, 62, 55, 68)
height
```

## Read it in as a `data.frame()`.

```{r dataframe}
mydf <- data.frame(id = 1:3,
                   group = c("a", "b", "c"),
                   names = c("Marielle", "Pablo", "Wyatt"))
mydf
```

## Using built in [to `R`] datasets.

```{r built_in}
# BUILT INTO R
data("USArrests")
str(USArrests)

# ACCESSED VIA PACKAGE
library(gapminder)
data("gapminder")
str(gapminder)
```

## Using external datasets

But... most usefully will be accessing external datasets. There are many formats 
of data - here are some common types and how to access them.

| Source Data | Base R Command | Tidyverse Package | 
| ----------- | -------------- | ----------------- | 
| CSV | `read.csv()` | `readr` | 
| Other delimitted text formats | `read.table()` | `readr` |
| Excel | | `readxl` |
| JSON | | `jsonlite` | 
| XML | | `xml2` |
| Web API | | `httr` |
| Scraping web pages | | `rvest` |

There are many many others, too, but we'll be using these (maybe).

### The `readr` Package

Though there are fully capable base R functions for the CSV data, we'll use 
the `readr` package in this class.

- It is faster (~10x faster)
- It returns `tibble`s
- More intuitive defaults
- Includes arguments for handing missing data [i.e. making `.` into a `NA`]

Functions in `readr`:

- `read_csv()` - comma separated files
- `read_csv2()` - semi-colon separated files
- `read_delim()` - generally delimited files (like a `|`)
- `read_fwf()` - fixed width files
- `read_log()` - application log files
- `read_table()` - space separated files
- `read_tsv()` - tab separated files

## Writing Data from R

Very often, you'll update a dataset or create a dataset and want to save it. 

In Base `R`, this is `write.csv()` or `write.table()`. 

In `readr`, this is `write_csv()` or `write_table()`.

### Our First Excel and CSVs File Reads

1. Download the `default_of_credit_card_clients.xlsx` dataset from the `dataset` 
folder on 
[Google drive](https://drive.google.com/drive/u/2/folders/1VmGLho1xioty-G7IaXCF2x_yKy-LPxC3). 
    - I recommend moving it into the same folder as your `Rmd` script.
2. Load the `readxl` package
    - It should be installed with the `tidyverse` package, but if not, use 
    `install.packages()` to install it.
3. Read in the file using the `read_xlsx()` function.
    - Assign the data to a variable called `credit`.
    - Verify that the data is there by using a `head()`, `str()`, and 
    `summary()`.
4. Using the `write_csv()` function, create a `.csv` version of your file with 
the same name in the same directory as the `.xlsx`.
    - Don't forget to load the `readr` package.
5. Read in the file using the `read_csv()` function.
    - Assign the data to a variable called `credit_csv`.
    - Verify that the data is there by using a `head()`, `str()`, and 
    `summary()`.

```{r in_class_together}
library(readxl)

credit <- read_xlsx("../../../Released to Students/datasets/default_of_credit_card_clients.xlsx")
head(credit)
str(credit)
summary(credit)

library(readr)

write_csv(credit, "../../../Released to Students/datasets/default_of_credit_card_clients.csv")

credit_csv <- read_csv("../../../Released to Students/datasets/default_of_credit_card_clients.csv")
head(credit_csv)
str(credit_csv)
summary(credit_csv)
```

## In Class Exercise

Let's do the in class exercise!