---
title: "Into to `dplyr`"
author: "Jake Rozran"
date: "2/15/2022"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(mdsr)
library(lubridate)

data("presidential")
head(presidential)
str(presidential)
```

# A grammar for Data Wrangling

With `ggplot2` we learned a grammar for data visualization. We were able to 
incrementally build data viz by linking commands (we did that with the `+`).

Now with `dplyr`, we use pipes (which look like `%>%`) to link data wrangling 
commands together. The first bunch:

- `select()`: Take a subset of columns
- `filter()`: Take a subset of rows
- `mutate()`: add a new column or modify an existing column
- `arrange()`: sort the rows
- `summarize()`: some aggregation across rows.
- `group_by()`: literally group by a feature in your data.

# Let's look at `select` and `filter`

```{r select}
select(presidential, name, party)
```

```{r filter}
filter(presidential, party == "Republican")
```

```{r filter_select_the_hard_way}
select(filter(presidential, party == "Republican"), name, party)
```

## Let's use `%>%` instead

```{r filter_with_pipe}
presidential %>% 
    filter(party == "Democratic")
```

```{r select_with_pipe}
presidential %>% 
    select(name, start)

```

```{r select_filter_atthesametime}
presidential %>% 
    filter(party == "Democratic") %>% 
    select(name, start)
```

# `mutate` and `rename`

```{r mutate}
presidential <- presidential %>% 
    mutate(term.length = interval(start, end) / dyears(1), 
           party = ifelse(party == "Republican", "R", "D")) %>% 
    rename(term_length = term.length)
presidential
```

# `arrange` function

```{r arrange}
presidential %>% 
    arrange(desc(term_length), desc(start))
```


# `summarize` and `group_by`

```{r summarise}
presidential %>% 
    group_by(party) %>% 
    summarise(num_years = sum(term_length),
              num_presidents = length(name),
              num_terms = num_years / 4) %>% 
    mutate(avg_stay = num_years / num_presidents) %>% 
    filter(party == "D") %>% 
    select(party, avg_stay)
```


















